---
title: "Data Preparation"
author: "Fazle Rabbi"
date: "October 21, 2015"
output: html_document
---

We now begin with the task of preparing our data for building models using R.

The required packages for this tutorial:
```{r}
library(rattle)                       # the weather dataset and normVarNames()
library(randomForest)                 # Input missing values using na. roughfix()
library(tidyr)                        # Tidy the dataset
library(ggplot2)                      # Visualize data
library(dplyr)                        # data preparation and pipes %>%
library(lubridate)                    # handle dates
library(FSelector)                    # feature selection
```

We begin by noting the path to the CSV file we wish to load
```{r}
dspath <- "http://rattle.togaware.com/weather.csv"
weather <- read.csv(dspath)
dsname <- "weather"
ds <- get(dsname)                # generic name
```

Once it is loaded we might like to get a basic idea of what it looks like
```{r}
dim(ds)
names(ds)
str(ds)
```

## Step 1: Convenience of Table Data Frame
Another tip in dealing with larger datasets is to make use of tbl df() to add a couple of extra
classes to the data frame. The simple aim here is to avoid the often made “mistake” of printing
the whole data frame accidentally.

```{r}
class(ds)
ds <- tbl_df(ds)
class(ds)
```

The default print() method reports the dimensions and then the first few rows and columns.
```{r}
ds
```

## Step 2: Review - Observations
```{r}
head(ds)
tail(ds)
ds[sample(nrow(ds),6),]    # sample creates random indices here (6 indices)
```

## Step 2: Review - Structure
```{r}
str(ds)
```

## Step 2: Review - Summay
The first quartile (Q1) is defined as the middle number between the smallest number and the median of the data set. 

```{r}
summary(ds)
```

## Step 2: Review - Meta Data Cleansing
**Normalize variable names**:  R is case sensitive. normVarNames() from rattle, converts variables into a standard form.
```{r}
names(ds) <- normVarNames(names(ds))
names(ds)
```






You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
