---
title: "data_modeling"
author: "Fazle Rabbi"
date: "October 25, 2015"
output: html_document
---

The packages that will be used for modeling are:

```{r}
library(rpart)          # Build decision tree model with rpart().
library(randomForest)   # Build model with randomForest().
library(ada)            # Build boosted trees model with ada().
library(rattle)         # Display tree model with fancyRpartPlot().
library(ROCR)           # Use prediction() for evaluation.
library(party)          # Build conditional tree models with ctree() and cforest().
library(ggplot2)        # Display evaluations.

```

## Step 4: Getting Started - Load the Dataset
In data preparation we saved the weather dataset, load it.

```{r}
(load("weather_151025.RData"))
dsname
dspath
dsdate
dim(ds)
id
target
risk
ignore
vars
```


## Step 4: Prepare - Formula to Describe the Goal
Predict rain_tomorrow from all other variables.

```{r}
(form <- formula(paste(target, "~ .")))  # target based on all other variables
```

## Step 4: Prepare - Training and Testing Datasets
Training, testing and validation datasets. First we initiate a random seed and report that seed so that the experiment can be repeated if necessary.

```{r}
(seed <- sample(1:1000000, 1))
seed <- 123
set.seed(seed)
```

Next we partition into training (70% random sample) and testing dataset.
```{r}
length(train <- sample(nobs, 0.7*nobs))
length(test  <- setdiff(seq_len(nobs), train))
```
The error rate calculated on testing dataset is closer to what we will obtain in general when we use the model. We also record the actual outcomes and the risks.

```{r}
actual.train <- ds[train, target]
actual <- ds[test, target]
risks <- ds[test, risk]
```

## Step 5: Build - Decision Tree
We build rpart() decision tree.

```{r}
ctrl <- rpart.control(maxdepth=3)
system.time(model <- m.rp <- rpart(form, ds[train, vars], control=ctrl))
model
mtype <- "rpart" # Record the type of the model for later use.

library(rpart.plot)
library(RColorBrewer)

fancyRpartPlot(model)
```


You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
